---
title: "Creating a new ggtheme"
author: "Jarred Robidoux"
date: "2023-02-07"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
library(ggplot2)
```

# __Building a New Theme__
Building and modifying a theme in ggplot2 is a key feature of the ggplot2 package and system for building data graphics. The ggplot2 package implements the notion of a theme for its plots by allowing you to modify many different elements of a plot and to store all those modifications as a special "theme" object. 

The default theme for ggplot2 is encapsulated by the theme_gray() function. Like other elements in the ggplot2 universe, themes can be "added" using the + operator to plot commands in order to change the look and feel of a plot. 

For example, here is a plot that uses the theme_classic() function: 
```{r}
knitr::opts_chunk$set(echo = TRUE)
mtcars %>%
  ggplot()+
  geom_point(mapping=aes(x=disp, y=mpg))+
  theme_classic()
```

# __Why Build a New Theme?__
For many people, building a new theme is a matter of personal preference with respect to colors, shapes, fonts, positioning of labels, etc. Because plots, much like writing, are an expression of your ideas, it is often desirable to customize those plots so that they accurately represent your vision. 

Developing themes is also a powerful branding tool. Plots that are distributed on the web or through marketing materials that have a common theme can be useful for reinforcing a brand. 

# __Default Theme__
As ntoed above, ggplot2 has a default theme, which is theme_gray(). This theme produces the familiar gray-background-white-grid-lines plot. You can obtain the default theme using the theme_get() function. 
```{r}
knitr::opts_chunk$set(echo = TRUE)
x <- theme_get()
class(x)
```
Notice that the object returned by theme_get() is an S3 object of class "theme" and "gg". This is the kind of object you will need to create or modify in order to customize your theme. 

You can modify the default theme by using the theme_set() function and passing it a theme object. For example, if we want all my plots to use the theme_minimal() theme, we could do
```{r}
knitr::opts_chunk$set(echo = TRUE)
new_theme <- theme_minimal()
theme_set(new_theme)
```
Now your plots will use the theme_minimal() theme without you having to specify it
```{r, echo = TRUE}
mtcars %>%
  ggplot()+
  geom_point(aes(x=disp, y=mpg))+
  facet_grid(.~gear)
```
Quitting R will erase the default theme setting. If you load ggplot2 in a future session it will revert to the default gray theme. If you'd like for ggplot2 to always use a different theme (either yours or one of the built-in ones), you can set a load hook and put it in your .Rprofile file. For example, the following hook sets the default theme to be theme_minimal() every time the ggplot2 package is loaded
```{r, echo = TRUE}
# setHook(packageEvent("ggplot2", "onLoad"),
        # function(...)ggplot2::theme_set(ggplot2::theme_minimal))
```

